---
title: "TIME-SERIES ANALYSIS- EXPLORATORY STUDY"
author: "MAHALAKSHMI"
output: 
  html_document: 
    keep_md: yes
---
<https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html>

```{r global_options, include=FALSE,cache=FALSE}
knitr::opts_chunk$set(echo = FALSE,echo=FALSE, message=FALSE, warning=FALSE,error=TRUE)
```

Reading the data and converting into a TS object

```{r kings}
kings <- scan("http://robjhyndman.com/tsdldata/misc/kings.dat",skip = 3)
kingsTS <- ts(kings)
births <- scan("http://robjhyndman.com/tsdldata/data/nybirths.dat")
birthsTS <- ts(births, frequency=12, start=c(1946,1))
souvenir <- scan("http://robjhyndman.com/tsdldata/data/fancy.dat")
souvenirTS <- ts(souvenir,frequency = 12,start = c(1987,1))
```

Plotting the read data

KingsTS can best be explianed by an additive model since the seasonal variations are random

```{r INITIAL PLOTS-1}
plot(kingsTS)

```

Birth plot can be explained best by an additive model since the seasonal variation is constant

```{r INITIAL PLOT-2}
plot(birthsTS)
```

The Plot of Souvenior has seasonal fluctuations and hence can be best explained by a multiplicative model (the amplitude of seasonal variance is not constant)

```{r INITIAL PLOT-3}
plot.ts(souvenirTS)

```

To convert a multiplicative model into a additive model take logarithms and the seasonality amplitude looks constant

```{r LOG Souvenir}
plot.ts(log(souvenirTS))

```



For a non seasonal data the trend can be explained by a simple moving average

```{r SMA}
library("TTR")

kingsTS3 <- SMA(kingsTS , n = 3)
plot(kingsTS3)

#smoothing it further
kingsTS9 <- SMA(kingsTS , n = 9)
plot(kingsTS9)

```


Decomposing : Means getting the trend/Seaonality and error

```{r DECOMPOSE ADDITIVE}
birthTSComponents <- decompose(birthsTS)
plot(birthTSComponents)

```

Decomposing a Multiplicative series- by subtracting the seasonality.Decompose is used to understand the trend without the effect of seasonality

```{r DECOMPOSE MULTIPLICATIVE}
souvenirComponents <- decompose(souvenirTS)
Souvenirnonseasonal <- souvenirTS - souvenirComponents$seasonal
plot(souvenirTS)
plot(Souvenirnonseasonal)
```

Simple Exponential smoothing : no trend and no seasonality

```{r Simple Expo smoothing}
rain <- scan("http://robjhyndman.com/tsdldata/hurst/precip1.dat",skip=1)
rainTS <- ts(rain,start=c(1813))
plot (rainTS)
```

### SIMPLE EXPONENTIAL SMOOTHING
Forecasting the above using HoltWinters.Simple exponential smoothing can be used for the above

```{r HW-1}
rainforecast <- HoltWinters(rainTS,beta = FALSE,gamma = FALSE, l.start = 23.56)
rainforecast
#Getting the forecast predicted
rainforecast$fitted
```

Plotting the forecast against the actual data

```{r PLOTTING FORECAST}
plot(rainforecast)
print (rainforecast$SSE)

```

Making forecast for the next 8 periods.

```{r FORECAST}
library(forecast)
rainforecast2 <- forecast(rainforecast,h=8)

```

Plotting the predicted forecast with a 80% and 95% confidence interval

```{r PLOTTING FORECAST}
plot(rainforecast2)

```

Getting the Correlogram of the residuals.If there is no corelation between the plotted acf's then the model cannot be improved.

```{r CORRELOGRAM-1}
acf(rainforecast2$residuals,lag.max = 20,na.action = na.omit)

```

As per the above the Linea t Lag -3 is just touching the significane boundary.This can be further checked using the Ljung-Box Test

```{r LJ-Box test}
Box.test(rainforecast2$residuals,lag = 20 , type = "Ljung-Box")

```

The p-value closer to 0 means there is a non-zero auto corealtion within the lags which is not the case in the above.Also the same can be checked by plotting the residuals to check for a zero mean and a constant variance

```{r RESIDUALS PLOT}
plot(rainforecast2$residuals)
```

## HOLTS EXPONENTIAL SMOOTHING

If there is trend and no seasonality then the above is used.

Trying out with the births data

```{r HOLTS EXPO}
skirts <- scan("http://robjhyndman.com/tsdldata/roberts/skirts.dat",skip=5)
skirtsTS <- ts(skirts,start=c(1866))
skirtsForecast <- HoltWinters(skirtsTS , gamma = FALSE)
print (skirtsForecast)

```

Plotting the forecast

```{r BIRTH FORECAST}
plot(skirtsForecast)


```

Making Forecast

```{r FORECASTING SKIRT}
skirtsForecast2 <- forecast(skirtsForecast,h=19)
plot(skirtsForecast2)
```

Testing the Possibility of betterment of the chosen method using ACF/Box-Lfung test and Plotting the residuals

```{r VALIDATION -1 }
acf(skirtsForecast2$residuals,lag = 20, na.action = na.omit)

# the lag at 5 is just chance .However using box method the same can be verified
Box.test(skirtsForecast2$residuals, lag = 20, type = "Ljung-Box")

#Plot showing randomness with mean 0 and constant variance
plot(skirtsForecast2$residuals)

```

## HOLT WINTERS EXPONENTIAL SMOOTHING

The parameters alpha, beta and gamma all have values between 0 and 1, and values that are close to 0 mean that relatively little weight is placed on the most recent observations when making forecasts of future values.

The Souvenir data looked like a multiplicative model beacuse of the seasonality rise.However the same can be converted to an additive model by taking Logarithms.

```{r LOG SOUVENIR}
souvenirlog <- log(souvenirTS)
souvenirforecast <- HoltWinters(souvenirlog)
print (souvenirforecast)
plot(souvenirforecast)
```

Forecasting for the next 48 months

```{r FORECASTING SOUVENIR}
souvenirforecast2  <- forecast(souvenirforecast, h=48)
plot(souvenirforecast2)

```

Checking the residual plot and the box method and ACF

```{r ACF,BOX and RESIDUAL PLOT}
plot(acf (souvenirforecast2$residuals,lag = 20, na.action = na.omit))
Box.test(souvenirforecast2$residuals,lag = 20, type = "Ljung-Box")
# P-value around 0.6
plot(souvenirforecast2$residuals)

```

Based on the above we can conclude that the model is adequate in predicting the forecast.

In cases where the residual plot does not have a zero mean and a constant variance ARIMA models are helpful.

Using the Skirt data

Differentiating the same to arrive at a stationary TS

```{r DIFF SKIRT}
skirtdiff1 <- diff(skirtsTS)
plot(skirtdiff1)

```


Differentiating once again since the above does not look stationary

```{r DIFF -2 SKIRT}
skirtdiff2 <- diff(skirtdiff1)
plot(skirtdiff2)

```

The above looks stationary hence we can proceed with the ARIMA(p,d,q) model.The order of differencing d for this case is 2.

```{r ACF and PACF}
plot(acf(skirtdiff2,lag = 20, plot = FALSE))
plot(pacf(skirtdiff2,lag = 20, plot = FALSE))

```

Both the ACF and PACF do not abruptly tail off and hence we can go for the ARIMA model(2,2,2).Also we can use the auto.ARIMA () to find the appropriate model

Another example

```{r KINGS ARIMA}
auto.arima(skirtsTS)
plot(kingsTS)
#differentiating to make it stationary
kingsdiff1 <- diff(kingsTS)
plot(kingsdiff1)

# d= 1
#Checking with ACF and PACf for Kings
plot (acf(kingsdiff1 , plot = FALSE,lag= 20))
plot (pacf(kingsdiff1 , plot = FALSE,lag= 20))

#ACF abrupt(q=1) and PACF gradual suggest MA model of order 1 
#ARIMA(0,1,1)
auto.arima(kingsTS)
```

Getting the ARIMA values and forecast

```{r KINGs ARIMA}

KingsARIMA <- arima(kingsTS, order = c(0,1,1))
kingsArimaForecast <- forecast(KingsARIMA , h= 10)
plot(kingsArimaForecast)

#checking with ACF ,PACF  and plotting the residuals
plot(acf(kingsArimaForecast$residuals, lag.max = 20 , plot = FALSE))
plot(pacf(kingsArimaForecast$residuals, lag.max = 20, plot = FALSE))
plot (kingsArimaForecast$residuals) # looks random
Box.test(kingsArimaForecast$residuals,lag = 20, type = "Ljung-Box")
plot(kingsArimaForecast)

```

Example of Volcanic dust Veil in Northern Hemisphere

```{r VOLCANO}
volcano <- scan("http://robjhyndman.com/tsdldata/annual/dvi.dat", skip=1)
volcanoTS <- ts(volcano)
plot(volcanoTS) # to check for diff
# No diff required d = 0
#Checking for P and q values
plot (acf(volcanoTS , lag = 20 , plot =FALSE))
plot (pacf(volcanoTS , lag = 20 , plot =FALSE))

#options AR(2,0) - based on PACF  and MA(0,3) based on ACF. but since ACF is gradual it is more an AR type with p =2.Checking with auto.arima

auto.arima(volcanoTS,ic= "bic")
#Best option ARIMA(2,0,0)
```
Forecasting using the suggested models

```{r FORECAST VOLCANO}
volcanoARIMA <- arima(volcanoTS, order = c(2,0,0))
volcanoforecast <- forecast(volcanoARIMA , h = 12)
plot(volcanoforecast)
plot(volcanoforecast$residuals)
plot(acf(volcanoforecast$residuals , lag = 20, plot = FALSE))
plot(pacf(volcanoforecast$residuals , lag = 20, plot = FALSE))
Box.test(volcanoforecast$residuals , type = "Ljung-Box")
```

