---
title: "REDWINE ANALYSIS"
author: "MAHALAKSHMI"
output: 
  html_document: 
    keep_md: yes
---

```{r global_options, include=FALSE,cache=FALSE}
knitr::opts_chunk$set(echo = FALSE,echo=FALSE, message=FALSE, warning=FALSE,error=TRUE)
```

###Loading the data 

The CSV file containing all the data points is loaded.
The data contains date points/Average temperature and the country names.

```{r LOADING}
temperature <- read.csv("C:/Users/Mash/Desktop/Git/Climate_TS/GlobalLandTemperaturesByCountry.csv")



```

### FILTERING THE DATA
Head function is used to understand the dataset.
The data pertaining to India is filtered using the dplyr package

```{r}
#head(temperature)
library(dplyr)
temp_only_IN <- temperature %>% filter(Country == "India") %>% select(AverageTemperature)
#print (temp_only_IN)
```

### CREATING A TS VARIABLE

Creating a TS variable using the filtered Indian temperatures using the TS 
command and setting the start and end accordingly.
A Further subset of the data from 1950 is considered for the prediction purpose.

```{r TS_INDIA}
TempTS <- ts(temp_only_IN,start=c(1796, 1), end=c(2013,9), frequency=12)
plot(TempTS)

#Subsetting and taking the data from 1950 only
TempTS2 <- window(TempTS, start=c(1950, 1), end=c(2014, 12))
library("zoo")
TempTS3 <-na.locf(TempTS2, option = "locf", na.remaining = "rev")

plot(TempTS3)
```

From the above plot  it can be seen that there is a seasonality component (which is constant) but not sure if there is a trend.Hence decomposing the same to find out the trend part.

```{r decompose}
temp_India_Components <- decompose(TempTS3)
plot(temp_India_Components)

```

The decompose  plot shows a clear trend(upward) and the seasonality remains constant.
We can go ahead with the Holt-Winters Exponential model with alpha,Beta nad Gamma parameters

```{r HWE}
TempForecast <- HoltWinters(TempTS3)
print (TempForecast$SSE)
#Predicting values for the next 72 months
library(forecast)
TempPrediction <- forecast(TempForecast, h = 72)
#print (TempPrediction$fitted)
plot(TempPrediction)
```

Let us check the residuals to find if they are any significant corelation using the acfa nd PACF plots.

```{r RESIDUALS}
plot(TempPrediction$residuals)

plot(acf(TempPrediction$residuals,plot= FALSE, na.action = na.omit))
plot(pacf(TempPrediction$residuals,plot= FALSE, na.action = na.omit))
```

Rechecking the co-relation using box test

```{r BOX_TEST}
Box.test(TempPrediction$residuals,lag = 20,type = "Ljung-Box")
```

As per box test there seems to be a corelation in the residuals , the above Holt-winters may not be appropriate and hence we can go for ARIMA models

```{r ARIMA}
#Checking for stationary before performing ARIMA
plot(TempTS3)
auto.arima(TempTS3)
```

The auto.Arima command suggest a 3,0,3 model.
The plot looks stationary hence we can gor for a ARIMA model with d=0
Checking for p and q values

```{r ACF_PACF}
plot(acf(TempTS3,lag = 100,plot = FALSE, na.action = na.omit))
plot(pacf(TempTS3,lag = 50,plot = FALSE, na.action = na.omit))
```

Applying ARIMA(2,0,0) and ARIMA(3,0,3) to get the best fit

```{r}
ARIMA1 <- arima(TempTS3 , order = c(2,0,0))
ARIMA2 <- arima(TempTS3, order = c(3,0,3))

ARIMA1pred <- forecast(ARIMA1, h = 48)
ARIMA2pred <- forecast(ARIMA2, h = 48)
```

Analysing both ARIMA1((2,0,0) and ARIMA2((3,0,3) as below

```{r ARIMA1_Validation}
plot(ARIMA1pred)
plot(ARIMA1pred$residuals)
Box.test(ARIMA1pred$residuals , type = "Ljung-Box")

#Plotting PACF and ACF

plot(acf(ARIMA1pred$residuals,lag = 100, plot= FALSE))
plot(pacf(ARIMA1pred$residuals,lag = 100, plot= FALSE))
```

Above does not seem to be the best choice as per the values in the Box - test and the plots

Checking for ARIMA2(3,0,3) as suggested by the auto function

```{r ARIMA2_VALIDATION}

plot(ARIMA2pred)
plot(ARIMA2pred$residuals)
Box.test(ARIMA2pred$residuals , type = "Ljung-Box")

#Plotting PACF and ACF

plot(acf(ARIMA2pred$residuals,lag = 100, plot= FALSE))
plot(pacf(ARIMA2pred$residuals,lag = 100, plot= FALSE))
```

The higher P-Value indiucates that actually there is no corelation between the residulas and hence the above model ARIMA(3,0,3) is the best choice and the prediction values are as below

```{r PREDICTED VALUES}
print (ARIMA2pred)

```

